---
title: "Tasks 1 & 2: Interactive plots"
toc: true
---

## Task 1: Session check-in

::: {.callout-tip title="Answer"}
Hopefully you wrote some interesting and muddy things!
:::


## Task 2: Interactive plot

The code below loads and cleans the data. **You don't need to modify anything in this chunk of code---you only need to run it.**

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(plotly)

# Technically this isn't necessary; you can run `penguins` and start using it 
# right away since it's a core dataset. This code just puts the dataset in the 
# environment panel so it's easier to see what's in there.
penguins <- datasets::penguins

# 11 rows are missing data on penguin sex, so we drop them
penguins <- penguins |> 
  drop_na(sex)
```

### Recreate plot

Use `ggplot()` and `ggplotly()` to recreate this interactive plot:

```{r}
#| label: recreate-me-interactive-plot
#| echo: false
#| message: false
#| warning: false
#| ref.label: recreation-interactive-plot
#| opts.label: true
```

**Some hints!**

- Don't try to make this interactive right away! [Just like in the example](https://datavizf25.classes.andrewheiss.com/example/11-example.html){target="_blank"}, first make a regular static version before feeding it into `ggplotly()`.
- The points are colored with three of R's built-in colors: `darkorange`, `purple`, and `cyan4`. Use `scale_color_manual()` to change the default colors to these (using something like `values = c("darkorange", "purple", "cyan4")`).
- The points are a little bigger than the default size. Use `size = BLAH` in `geom_point()` to adjust them (but make sure you don't use that inside `aes()`!).
- Notice that the points use different shapes for each species as well as color!
- The plot uses `theme_light()`.
- The interactive tooltip should be structured like this:

  ```default
  Species (sex on island)
  Bill depth: X mm
  Bill length: X mm
  ```

  You'll need to make a new column in the penguins dataset named `label` (or whatever you want to call it) that builds that label using that template. [The example will be useful](https://datavizf25.classes.andrewheiss.com/example/11-example.html#including-more-information-in-the-tooltip){target="_blank"}.

  ::: {.callout-tip collapse="true"}
  ### Bonus tip!

  In the example, I use `paste0()` to build the label text, like this:

  ```{r}
  penguins |> 
    mutate(fancy_label = paste0(species, " (", sex, " on ", island, " Island)")) |> 
    select(fancy_label) |> head(2)
  ```

  …and that works, but it's *super annoying* to switch between variable names and text and to keep track of where spaces need to be in the text.

  To make life easier, try out [the {glue} package](https://glue.tidyverse.org/){target="_blank"}, which is installed as part of the tidyverse, but not loaded with `library(tidyverse)`. {glue} lets you substitute variable values directly in text without needing to separate everything with commas. Anything inside curly braces `{}` will get replaced with the value in the data:

  ```{r}
  library(glue)

  penguins |> 
    mutate(fancy_label = glue("{species} ({sex} on {island} Island)")) |> 
    select(fancy_label) |> head(2)
  ```

  :::

### Answer

::: {.callout-tip title="Answer"}

```{r}
#| label: recreation-interactive-plot
#| warning: false
#| echo: true

penguins_labeled <- penguins |> 
  mutate(label = glue::glue(
    "{species} ({sex} on {island} Island)<br>Bill depth: {bill_dep}<br>Bill length: {bill_len}"
  ))

p <- ggplot(penguins_labeled, aes(x = bill_dep, y = bill_len)) +
  geom_point(size = 2, aes(text = label, color = species, shape = species)) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  theme_light()

ggplotly(p, tooltip = "text")
```

:::
